---
title: "Friends Dialogue"
description: |
  Analysis of Friends Dialogue
author: Mayra Coruh
date: October 2, 2024
format: html
execute:
  warning: false
  message: false
---

Source of Data: https://emilhvitfeldt.github.io/friends/

```{r}
library(friends)
library(stringr)
library(tidyr)
library(dplyr)
library(ggplot2)
data(friends)
```

# Finding the number of times Rachel says "honey" across all seasons

```{r}
rachel_dialogue <- friends |>
  dplyr::filter(speaker == "Rachel Green") |>
  select(season, text)

rachel_honey_count <- rachel_dialogue |>
  mutate(honey_count = str_count(text, "(?i)\\bhoney\\b"))

honey_by_season <- rachel_honey_count |>
  group_by(season) |> 
  summarise(total_honey = sum(honey_count))

ggplot(honey_by_season, aes(x = season, y = total_honey)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  theme_minimal() +
  labs(
      title = "Number of Times Rachel Says 'Honey' per Season",
       x = "Season",
       y = "Number of 'Honey' Occurrences"
)

```

# Description of Insights

The histogram illustrates how the use of the word "honey" by Rachel fluctuates across seasons, with some seasons showing higher frequencies than others. Rachel's use of "honey" peaks during Season 3 with the highest number of occurrence of 40. Another peak occurs in Season 5, with a number of occurrence of approximately 35. In seasons 2 and 7, there is a moderate use of the word "honey," with occurrences around 20-25, suggesting that Rachel uses "honey" consistently but less frequently in these seasons. The lowest number of times "honey" is used is in Season 1, with less than 10 occurrences. Similarly, Rachel uses "honey" less in later seasons. For instance, in Season 10, number of occurrences falls below 10. Overall, the plot indicates that Rachel's use of the term "honey" peaks during the middle seasons of the the show and drops off significantly in the early and final seasons. I wonder how this might relate to a more nuanced analysis on the relationship between Rachel's emotional state and dialogue across seasons. So, for instance, can there be a correlation between her use of "honey" and her emotional state? This analysis could be interesting if particular events in her life are taken as reference points, such as her relationship with Ross (times/seasons when their together and when they are not etc.).

# Analysis of Utterance Categories

```{r}
summary_data <- friends |>
  mutate(text_lower = str_to_lower(text),  
        is_love = str_detect(text_lower, "\\blove\\b"),
        is_question = str_detect(text_lower, "^(what|where|who|when|how)"),
        has_multiple_exclamations = str_detect(text_lower, "!{2,}")
  )

summary_data <- summary_data |>
  summarise(
      love_statements = sum(is_love),
      questions = sum(is_question),
      multiple_exclamations = sum(has_multiple_exclamations)
      ) |>
  pivot_longer(cols = everything(), names_to = "category", values_to = "count")

 # total_data = full_join(summary_data, friends, by ="speaker") --> new fata frame that has count and category columns. the problem is that the category is not in summary_data

summary_data |>
  ggplot(aes(x = category, y = count, fill = category)) +
  geom_bar(stat = "identity") +
  labs(title = "Analysis of Utterance Categories", x = "Category", y = "Count") +
  theme_minimal()
```

# Description of Insights

The histogram provides information on the different types of utterances across three categories: love statements, multiple exclamations, and questions. The plot depicts that the questions category has the highest count of occurrences with over 3000 instances, indicating that speakers in the show ask questions more frequently than they express love or use exclamations. Furthermore, having a count between 1500 and 2000, multiple_exclamations is the second most frequent category. This suggests that expressions of excitement or heightened emotions, which are identified by multiple exclamations, are relatively common but not as common as questions. Lastly, love statements is the least frequent category, with a count of around 1000. This implies that expressions of love occur less frequently compared to the other two categories. Overall, the graph analyzes the nature of communication in Friends across the three parameters of love statements, multiple exclamations, and questions.
