---
title: "SQL"
author: Mayra Coruh
date: November 26, 2024
format: html
execute:
  warning: false
  message: false
---

```{r}
library(mdsr)
library(dbplyr)
library(DBI)
library(RMariaDB)
library(tidyverse)
```

Source of Data: Wideband Acoustic Immittance (WAI) Database hosted by Smith College (doi.org/10.35482/egr.001.2022)

Plan:

In this project, I plan to analyze and visualize data from the Wideband Acoustic Immittance (WAI) Database. To do so, I will consider the absorbance measurements across different frequencies for specific identifiers and groups. I will explore patterns in absorbance by using SQL queries, for instance, by grouping data by variables such as sex and frequency. Lastly, using ggplot, I intend to visualize the processed data in SQL to visually represent the explored relationships between variables.

```{r}
con_wai <- dbConnect(
  MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname = "wai"
)
Measurements <- tbl(con_wai, "Measurements")
PI_Info <- tbl(con_wai, "PI_Info")
Subjects <- tbl(con_wai, "Subjects")

# collect(Measurements)
```


```{sql}

#| connection: con_wai

# Step 1: Explore the structure of the WAI database by listing available tables
# This provides an overview of the database's components.

SHOW TABLES;
```

```{sql}

#| connection: con_wai

# Step 2: Preview the first few rows of the 'Measurements' table
# This allows us to understand the structure and key variables of the dataset.

SELECT *
FROM Measurements
LIMIT 0, 5;
```

```{sql}

#| connection: con_wai

# Step 3: Filter detailed information for a specific study (Abur_2014)
# The 'PI_Info' table contains metadata about studies in the database.

SELECT * FROM PI_Info 
WHERE Identifier = "Abur_2014"; 
```

```{sql}

#| connection: con_wai

# Step 4: Filter absorbance measurements for the study 'Abur_2014'
# This query focuses on analyzing measurement data for the chosen study.

SELECT * FROM Measurements 
WHERE Identifier = "Abur_2014"; 
```

```{sql}

#| connection: con_wai
#| output.var: "graph"


# Step 5: Aggregate mean absorbance across frequencies for selected studies
# This query filters data for specific studies and calculates mean absorbance 
# for each frequency within the range 200-8000 Hz.
# The results will be visualized later to explore absorbance patterns.

SELECT Identifier, Frequency, LOG10(Frequency) AS log_frequency, AVG(Absorbance) AS mean_absorbance 
FROM Measurements
WHERE Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", "Liu_2008"
"Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000
GROUP BY Identifier, Frequency;


```

```{r}

# Step 6: Visualize mean absorbance across frequencies for various studies
# This plot depicts how mean absorbance changes with frequency for selected studies.

graph |>
ggplot(aes (x = Frequency, y = mean_absorbance, 
  color = Identifier,
  group = Identifier)) +
  geom_line()+
  scale_x_log10()
  labs(
    title = "Mean Absorbance Across Frequencies for Various Studies",
    x = "Frequency (Log Scale)",
    y = "Mean Absorbance"
  )

```

Description of Graph:

This plot titled "Mean Absorbance Across Frequencies for Various Studies" shows the mean absorbance values across a range of frequencies (plotted on a log scale) for various studies, as identified by the color-coded legend. The graph, where each line represents data from a specific study, depicts how absorbance varies with frequency within the range of 200 to 8000. Overall, the absorbance increases with frequency, reaching a peak between approximately 1000 and 3000 before declining.

```{sql}

#| connection: con_wai

# Step 7: Join PI_Info and Measurements tables to count unique ears measured
# This query provides additional context about the dataset by linking study metadata
# with measurement data and counting distinct subject-ear combinations.

SELECT p.Identifier, p.Year, p.AuthorsShortList, 
 COUNT(DISTINCT SubjectNumber, Ear) AS ear_u
FROM PI_Info AS p 
LEFT JOIN Measurements AS m ON m.Identifier = p.Identifier 
WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", "Liu_2008"
"Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000
GROUP BY Identifier, Instrument;
```

```{sql}

#| connection: con_wai

# Step 8: Generate legend strings for studies with metadata and ear counts
# This query creates formatted legend labels that combine study metadata 
# (authors, year, sample size, instrument) to be used in visualization.

SELECT p.Identifier, Year, AuthorsShortList, 
 COUNT(DISTINCT SubjectNumber, Ear) AS ear_u,
 CONCAT(AuthorsShortList, " (" , year, ") ", "N=", COUNT(DISTINCT SubjectNumber, Ear), "; ", Instrument) AS legend
FROM PI_Info AS p 
LEFT JOIN Measurements AS m ON m.Identifier = p.Identifier 
WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", "Liu_2008"
"Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000
GROUP BY Identifier, Instrument;

```

```{sql}

#| connection: con_wai
#| output.var: "graph2"


# Step 9: Prepare data for graph visualization
# This query extends the previous aggregation by including study metadata,
# calculating mean absorbance across frequencies, and formatting legend labels.

SELECT p.Identifier, p.Year, p.AuthorsShortList, Frequency,
 LOG10(Frequency) AS log_frequency, AVG(Absorbance) AS mean_absorbance,
 COUNT(DISTINCT SubjectNumber, Ear) AS ear_u,
 CONCAT(AuthorsShortList, " (" , year, ") ", "N=", COUNT(DISTINCT SubjectNumber, Ear), "; ", Instrument) AS legend
FROM PI_Info AS p 
LEFT JOIN Measurements AS m ON m.Identifier = p.Identifier 
WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015" ,"Lewis_2015", "Liu_2008"
"Rosowski_2012", "Shahnaz_2006", "Shaver_2013" , "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010" ) AND Frequency > 200 AND Frequency < 8000
GROUP BY Identifier, Instrument, Frequency;
```

```{sql}

#| connection: con_wai

# Step 10: Explore subject metadata
# This query lists all entries in the Subjects table to inspect participant information.

SELECT * 
FROM Subjects
```

```{sql}

#| connection: con_wai
#| output.var: "graphforQ2"

# Step 11: Analyze absorbance by sex for a specific study (Lewis_2015)
# This query calculates mean absorbance for males and females separately 
# for each frequency within the specified range.

SELECT Sex, Frequency, 
    AVG(Absorbance) AS mean_absorbance, m.Identifier
FROM Subjects AS s
RIGHT JOIN Measurements AS m ON m.Identifier = s.Identifier 
WHERE m.Identifier = "Lewis_2015" AND Frequency > 200 AND Frequency < 8000
GROUP BY m.Identifier, Frequency, Sex;
```

```{r}

# Step 12: Display the first few rows of the processed data for inspection

head(graphforQ2)
```

```{r}

# Step 13: Visualize absorbance trends for females in the 'Lewis_2015' study
# This plot focuses on female participants, showing mean absorbance across frequencies.
graphforQ2 |>
  filter(Sex == "Female") |>
  ggplot(aes(
    x = Frequency,
    y = mean_absorbance,
    color = Sex,
    group = Sex
  )) +
  geom_line() +
  scale_x_log10() +
  labs(
    title = "Mean Absorbance Across Frequencies for Females", 
    x = "Frequency",
    y = "mean_absorbance", 
    color = "Identifier"  
  )


```

Description of Plot:

This plot displays the mean absorbance across a range of frequencies (plotted on a log scale) for female participants. The red line represents how the absorbance changes as frequency increases, starting from around 300 and extending up to approximately 8000. Absorbance increases steadily at lower frequencies, reaching a peak between 1000 and 3000 before declining at higher frequencies.

Description of Code:

In the code above, I, initially, connected to the WAI database and used SQL queries to extract absorbance data grouped by identifiers, sex, and frequency, applying necessary filters for frequency range and study identifiers. Then, piping in the processed data into R, I visualized it via ggplot. For example, I plotted mean absorbance by frequency on a log scale, distinguishing data by attributes like sex or study identifiers.
